---
title: "NuVasive Mergers & Acquisitions Event Study"
author: "Greg Robinson"
date: "September 26, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### EXECUTIVE SUMMARY

I conduct a NuVasive-focused mergers and acquisitions (M&A) event study analysis between May 13, 2004 and Dec. 31, 2016 using the Thomson-Reuters SDC Platinum M&A database, security and index returns from the Center for Research in Security Prices (CRSP), and Fama-French empirical asset pricing factor returns.  I show that NuVasive as a whole outperforms its industry (Standard Industry Classification code 3841 - Surgical and Medical Instruments and Apparatus) and the market based on a 5-day Cumulative Abnormal Return (CAR) around the acquisition announcement date. A portfolio of all acquiring firms (excluding financial services and real estate) underperforms NuVasive in the short-run and outperforms NuVasive in the long-run based on a 36-month Buy-and-Hold Abnormal Return (BHAR).  


#### METHODOLOGY & THEORY

Mergers and acquisitions event studies are based on identifying abnormal security returns for acquiring and target firms around the acquisition announcement date, where abnormal returns are the difference between actual and predicted returns at date t:

* Abnormal Returns = Actual Returns - Predicted Returns



\[
AR_{it} = R_{it} - \hat{R_{it}}
\]  

Significance of the abnormal return is then calculated based on a test statistic:

\[
z_{it} = \frac{AR_{it}}{\sigma_{AR_{it}}}
\]  

**Cumulative Abnormal Returns (CARs):**  I use an estimation period of [-242, -42] days prior to the acquisition announcement date to generate the counterfactual predicted returns for both acquiring and target firms if the acquisition did not take place.  I then calculate abnormal returns against a benchmark portfolio during the event window of [-2, 2] days around the acquisition announcement date.  CARs identify information leakage prior to the public announcement, as well as immediate and delayed market reaction to the event.

\[
CAR_{it} = \sum_{t=-2}^{t=2}AR_{it}
\]  

**Buy-and-Hold Abnormal Returns (BHARs):**  I use an event period of 36 months following the acquisition announcement date to generate long-term abnormal security returns against a benchmark.  BHARs are meant to replicate the experience of an actual investor, and reflect long-term firm performance following an event.

\[
BHAR_{it} = \prod_{t=0}^{t=36}(1+R_{it}) - \prod_{t=0}^{t=36}(1+R_{mt})
\]  

In practice, BHARs are calculated as:

\[
BHAR_{it} = e^\sum{log(1+R_{it})} - e^\sum{log(1+R_{mt})}
\]  

where $R_{mt}$ is the benchmark portfolio return.

**Benchmarks:**  Empirical asset pricing models used to generate expected returns.

- **Market Model:** Capital Asset Pricing Model (CAPM) that describes the relationship between systematic risk and expected return for stocks

\[
 (\hat{R_{it}} - rf) = \hat{\alpha} + \hat{\beta_{i}}(R_{mt} - rf) + \epsilon
\]  

+ where $R_{mt}$ is the expected return of the risky security (acquiring or target firms), $rf$  is the risk-free rate, and $R_{mt}$ represents the return of the market portfolio. The Market Model is used for calculating both CARs and BHARs in my analysis.

- **Fama-French 4-Factor Model:** Empirical asset pricing model based on the Market Factor (CAPM Beta) to represent undiversifiable risk, the Size Factor (SMB) to reflect small firms historically outperforming large firms, the Value Factor (HML) to reflect value firms historically outperforming growth firms, and the Momentum Factor (MOM) to reflect the tendency of securities that have outperformed or underperformed the market in a recent period to continue to do so.  The FF-4 model is calculated using both equal and value-weighted market portfolios, and is only used for calculating portfolio BHARs in my analysis.

\[
 (\hat{R_{it}}- rf) = \hat{\alpha}+\hat{\beta_{1}}(R_{mt} - rf) + \hat{\beta_{2}}(SMB) + \hat{\beta_{3}}(HML) + \hat{\beta_{4}}(MOM) + \epsilon
\]  

#### ANALYSIS

**All Firms (excluding Financial Services and Real Estate, SIC 60-67)**

I begin with an analysis of all firms, excluding financial services and real estate based on SIC 60-67.   Acquiring firms have slightly negative CARs around the acquisition announcement period while Target firms have highly positive CARs.  Over a 36-month period, however, acquiring firms have significant BHARs.


```{r, echo = FALSE}
library(knitr)
library(ggplot2)

All <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/MD_allcleansun_fin.csv", header = TRUE, stringsAsFactors = FALSE)
All$DateAnnounced <- as.Date(as.character(All$DateAnnounced), format="%m/%d/%Y")
All$Year <- substr(All$DateAnnounced, 1, 4)
All$Year <- as.factor(All$Year)


Hist1 <- ggplot(All, aes(Year)) +
  geom_bar() +
  labs(x="Year", y="# of M&A transactions") +
  ggtitle("M&A Frequency: All Firms")

print(Hist1)
```

```{r echo=FALSE, message=FALSE, results='asis', comment=""}

library(knitr)

t1.1 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/Table1.csv", header = TRUE, stringsAsFactors = FALSE)

names(t1.1)[1] <- "Abnormal Return Metric"
names(t1.1)[4] <- "Std. Err."
names(t1.1)[5] <- "95% Conf."
names(t1.1)[6] <- "Interval"

kable(t1.1, caption = "All Firms")
```

Triple asterisk indicates significance at the 1% confidence level

**All Companies operating under any SIC code 3841 (Surgical and Medical Instruments and Apparatus)**

I restrict the analysis to all companies operating under any SIC code 3841 (primary, secondary, etc.) while excluding NuVasive in order to isolate its individual performance.  This general group includes NuVasive’s industry, as well as unrelated acquisitions by firms operating in multiple business segments that include SIC 3841.  Acquirer CARs and BHARs are not statistically different from zero based on either benchmark, while Target firms significantly outperform the market portfolio by 41 percent during the event window [-2, 2].

```{r, echo = FALSE}
library(knitr)
library(ggplot2)

All1 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/Md_All3841SUNDAY.csv", header = TRUE, stringsAsFactors = FALSE)
All1$DateAnnounced <- as.Date(as.character(All1$DateAnnounced), format="%m/%d/%Y")
All1$Year <- substr(All1$DateAnnounced, 1, 4)
All1$Year <- as.factor(All1$Year)


Hist2 <- ggplot(All1, aes(Year)) +
  geom_bar() +
  labs(x="Year", y="# of M&A transactions") +
  ggtitle("M&A Frequency: All SIC 3841")

print(Hist2)
```

```{r echo=FALSE, message=FALSE, results='asis', comment=""}

library(knitr)

t2 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/Table2.csv", header = TRUE, stringsAsFactors = FALSE)

names(t2)[1] <- "Abnormal Return Metric"
names(t2)[4] <- "Std. Err."
names(t2)[5] <- "95% Conf."
names(t2)[6] <- "Interval"

kable(t2, caption = "All Firms with SIC 3841")
```

Triple asterisk indicates significance at the 1% confidence level

**Companies operating under Primary SIC code 3841 only**

I further restrict the analysis to firms operating under a primary SIC 3841 only, again excluding NuVasive in order to isolate its performance.  This group is the best representation of NuVasive’s industry and competitors.  Acquirer CARs and BHARs based on either benchmark are not statistically different from zero, while target CARs outperform the market portfolio by 28 percent around the event window. 

```{r, echo = FALSE}
library(knitr)
library(ggplot2)

All3 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/MD_AllPrime3841SUNDAY.csv", header = TRUE, stringsAsFactors = FALSE)
All3$DateAnnounced <- as.Date(as.character(All3$DateAnnounced), format="%m/%d/%Y")
All3$Year <- substr(All3$DateAnnounced, 1, 4)
All3$Year <- as.factor(All3$Year)


Hist3 <- ggplot(All3, aes(Year)) +
  geom_bar() +
  labs(x="Year", y="# of M&A transactions") +
  ggtitle("M&A Frequency: Primary SIC 3841 Only")

print(Hist3)
```

```{r echo=FALSE, message=FALSE, results='asis', comment=""}

library(knitr)

t3 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/Table3.csv", header = TRUE, stringsAsFactors = FALSE)

names(t3)[1] <- "Abnormal Return Metric"
names(t3)[4] <- "Std. Err."
names(t3)[5] <- "95% Conf."
names(t3)[6] <- "Interval"

kable(t3, caption = "Firms with Primary SIC 3841 Only")
```


####All NuVasive Combined Acquisitions

I combine all NuVasive-only acquisitions for which there is available data into a portfolio in order to evaluate the company’s cumulative post-acquisition performance since its IPO date.  All NuVasive target firms were private.  NuVasive outperforms the market portfolio by 4.5 percent around the event window at a statistically significant level.

```{r echo=FALSE, message=FALSE, results='asis', comment="", warning = FALSE}
library(knitr)

t4 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/AllNuvAcqData1.csv", header = TRUE, stringsAsFactors = FALSE)

names(t4)[1] <- "Abnormal Return Metric"
names(t4)[4] <- "Std. Err."
names(t4)[5] <- "95% Conf."
names(t4)[6] <- "Interval"

kable(t4, caption = "All NuVasive (combined) Acquisitions")
```

Single asterisk indicates omission of one acquisition due to insufficient data; double asterisk indicates significance at the 10% confidence level


####NuVasive Individual Acquisitions

I examine each of NuVasive’s individual acquisitions based on CARs, BHARs (when there is sufficient available data), and graphically.

```{r echo=FALSE, message=FALSE, results='asis', comment="", warning = FALSE}
library(knitr)

t5 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/IndAcqTable.csv", header = TRUE, stringsAsFactors = FALSE)

names(t5)[1] <- "Target Name"
names(t5)[2] <- "Announcement Date"
names(t5)[3] <- "Effective Date"
names(t5)[4] <- "Perc. Shares Acq."
names(t5)[5] <- "CARs"
names(t5)[6] <- "BHARs"

kable(t5, caption = "Individual NuVasive Acquisitions")
```

**RSB Spine – Cervical Plate Acquisition**

```{r, echo = FALSE, warnings = FALSE}
library(knitr)
library(ggplot2)
library(gridExtra)

ACAR20050606 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/ACAR20050606.csv", header = TRUE, stringsAsFactors = FALSE)
ACAR20050606$DATE <- as.Date(as.character(ACAR20050606$DATE), format="%Y%m%d")
ACAR20050606$DATE <- as.character(ACAR20050606$DATE)

ACAR20050606G <- ggplot(ACAR20050606) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=AbnRet), group=1, color = 'red') +
  labs(title = "RSB Spine: 5-Day Cumulative Abnormal Returns (CARs)", subtitle="Market return (Black), NuVasive return Blue), NuVasive 5-Day CARs (Red)",
       x ="Acquisition Announcement Date: 2005-06-06", y="% Return")
  
  
BHAR20050606 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/BHAR20050606.csv", header = TRUE, stringsAsFactors = FALSE)
BHAR20050606$DATE <- as.Date(as.character(BHAR20050606$DATE), format="%Y%m%d")
#BHAR20050606$DATE <- as.character(BHAR20050606$DATE)
BHAR20050606$BHAR <- (exp(cumsum(log(1+BHAR20050606$ARET)))) - (exp(cumsum(log(1+BHAR20050606$mret))))

BHAR20050606G <- ggplot(BHAR20050606) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=BHAR), group=1, color = 'red') +
  labs(title = "RSB Spine: 36-Month Buy-and-Hold Abnormal Returns (BHARs)", subtitle="Market return (Black), NuVasive return, (Blue), NuVasive 36-Mon BHARs (Red)",
       x ="Acquisition Announcement Date: 2005-06-06", y="% Return")


First <- grid.arrange(ACAR20050606G, BHAR20050606G, ncol=1)
```

**RiverBend Design Acquisition**

```{r, echo = FALSE, warnings = FALSE}
library(knitr)
library(ggplot2)
library(gridExtra)

ACAR20050812 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/ACAR20050812.csv", header = TRUE, stringsAsFactors = FALSE)
ACAR20050812$DATE <- as.Date(as.character(ACAR20050812$DATE), format="%Y%m%d")
ACAR20050812$DATE <- as.character(ACAR20050812$DATE)

ACAR20050812G <- ggplot(ACAR20050812) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=AbnRet), group=1, color = 'red') +
  labs(title = "RiverBend Designs: 5-Day Cumulative Abnormal Returns (CARs)", subtitle="Market return (Black), NuVasive return Blue), NuVasive 5-Day CARs (Red)",
       x ="Acquisition Announcement Date: 2005-08-12", y="% Return")

BHAR20050812 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/BHAR20050812.csv", header = TRUE, stringsAsFactors = FALSE)
BHAR20050812$DATE <- as.Date(as.character(BHAR20050812$DATE), format="%Y%m%d")
#BHAR20050812$DATE <- as.character(BHAR20050812$DATE)
BHAR20050812$BHAR <- (exp(cumsum(log(1+BHAR20050812$ARET)))) - (exp(cumsum(log(1+BHAR20050812$mret))))

BHAR20050812G <- ggplot(BHAR20050812) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=BHAR), group=1, color = 'red') +
  labs(title = "RiverBend Designs: 36-Month Buy-and-Hold Abnormal Returns (BHARs)", subtitle="Market return (Black), NuVasive return, (Blue), NuVasive 36-Mon BHARs (Red)",
       x ="Acquisition Announcement Date: 2005-06-06", y="% Return")


Second <- grid.arrange(ACAR20050812G, BHAR20050812G, ncol=1)
```

**Radius Medical Acquisition**

```{r, echo = FALSE, warnings = FALSE}
library(knitr)
library(ggplot2)
library(gridExtra)

ACAR20070123 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/ACAR20070123.csv", header = TRUE, stringsAsFactors = FALSE)
ACAR20070123$DATE <- as.Date(as.character(ACAR20070123$DATE), format="%Y%m%d")
ACAR20070123$DATE <- as.character(ACAR20070123$DATE)

ACAR20070123G <- ggplot(ACAR20070123) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=AbnRet), group=1, color = 'red') +
  labs(title = "Radius Medical: 5-Day Cumulative Abnormal Returns (CARs)", subtitle="Market return (Black), NuVasive return Blue), NuVasive 5-Day CARs (Red)",
       x ="Acquisition Announcement Date: 2007-01-23", y="% Return")

BHAR20070123 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/BHAR20070123.csv", header = TRUE, stringsAsFactors = FALSE)
BHAR20070123$DATE <- as.Date(as.character(BHAR20070123$DATE), format="%Y%m%d")
#BHAR20070123$DATE <- as.character(BHAR20070123$DATE)
BHAR20070123$BHAR <- (exp(cumsum(log(1+BHAR20070123$ARET)))) - (exp(cumsum(log(1+BHAR20070123$mret))))

BHAR20070123G <- ggplot(BHAR20070123) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=BHAR), group=1, color = 'red') +
  labs(title = "Radius Medical: 36-Month Buy-and-Hold Abnormal Returns (BHARs)", subtitle="Market return (Black), NuVasive return, (Blue), NuVasive 36-Mon BHARs (Red)",
       x ="Acquisition Announcement Date: 2007-01-2", y="% Return")


Third <- grid.arrange(ACAR20070123G, BHAR20070123G, ncol=1)
```

**Osiris Therapeutics Acquisition**

```{r, echo = FALSE, warnings = FALSE}
library(knitr)
library(ggplot2)
library(gridExtra)

ACAR20080508 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/ACAR20080508.csv", header = TRUE, stringsAsFactors = FALSE)
ACAR20080508$DATE <- as.Date(as.character(ACAR20080508$DATE), format="%Y%m%d")
ACAR20080508$DATE <- as.character(ACAR20080508$DATE)

ACAR20080508G <- ggplot(ACAR20080508) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=AbnRet), group=1, color = 'red') +
  labs(title = "Osiris Therapeutics: 5-Day Cumulative Abnormal Returns (CARs)", subtitle="Market return (Black), NuVasive return Blue), NuVasive 5-Day CARs (Red)",
       x ="Acquisition Announcement Date: 2008-05-08", y="% Return")

BHAR20080508 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/BHAR20080508.csv", header = TRUE, stringsAsFactors = FALSE)
BHAR20080508$DATE <- as.Date(as.character(BHAR20080508$DATE), format="%Y%m%d")
#BHAR20080508$DATE <- as.character(BHAR20080508$DATE)
BHAR20080508$BHAR <- (exp(cumsum(log(1+BHAR20080508$ARET)))) - (exp(cumsum(log(1+BHAR20080508$mret))))

BHAR20080508G <- ggplot(BHAR20080508) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=BHAR), group=1, color = 'red') +
  labs(title = "Osiris Therapeutics: 36-Month Buy-and-Hold Abnormal Returns (BHARs)", subtitle="Market return (Black), NuVasive return, (Blue), NuVasive 36-Mon BHARs (Red)",
       x ="Acquisition Announcement Date: 2008-05-08", y="% Return")


Fourth <- grid.arrange(ACAR20080508G, BHAR20080508G, ncol=1)
```

**Cervitech Acquisition**

```{r, echo = FALSE, warnings = FALSE}
library(knitr)
library(ggplot2)
library(gridExtra)

ACAR20090422 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/ACAR20090422.csv", header = TRUE, stringsAsFactors = FALSE)
ACAR20090422$DATE <- as.Date(as.character(ACAR20090422$DATE), format="%Y%m%d")
ACAR20090422$DATE <- as.character(ACAR20090422$DATE)

ACAR20090422G <- ggplot(ACAR20090422) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=AbnRet), group=1, color = 'red') +
  labs(title = "Cervitech: 5-Day Cumulative Abnormal Returns (CARs)", subtitle="Market return (Black), NuVasive return Blue), NuVasive 5-Day CARs (Red)",
       x ="Acquisition Announcement Date: 2009-04-22", y="% Return")

BHAR20090422 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/BHAR20090422.csv", header = TRUE, stringsAsFactors = FALSE)
BHAR20090422$DATE <- as.Date(as.character(BHAR20090422$DATE), format="%Y%m%d")
#BHAR20090422$DATE <- as.character(BHAR20090422$DATE)
BHAR20090422$BHAR <- (exp(cumsum(log(1+BHAR20090422$ARET)))) - (exp(cumsum(log(1+BHAR20090422$mret))))

BHAR20090422G <- ggplot(BHAR20090422) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=BHAR), group=1, color = 'red') +
  labs(title = "Cervitech: 36-Month Buy-and-Hold Abnormal Returns (BHARs)", subtitle="Market return (Black), NuVasive return, (Blue), NuVasive 36-Mon BHARs (Red)",
       x ="Acquisition Announcement Date: 2009-04-22", y="% Return")


Fifth <- grid.arrange(ACAR20090422G, BHAR20090422G, ncol=1)
```

**ANC Acquisition**

```{r, echo = FALSE, warnings = FALSE}
library(knitr)
library(ggplot2)
library(gridExtra)

ACAR20130506 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/ACAR20130506.csv", header = TRUE, stringsAsFactors = FALSE)
ACAR20130506$DATE <- as.Date(as.character(ACAR20130506$DATE), format="%Y%m%d")
ACAR20130506$DATE <- as.character(ACAR20130506$DATE)

ACAR20130506G <- ggplot(ACAR20130506) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=AbnRet), group=1, color = 'red') +
  labs(title = "ANC: 5-Day Cumulative Abnormal Returns (CARs)", subtitle="Market return (Black), NuVasive return Blue), NuVasive 5-Day CARs (Red)",
       x ="Acquisition Announcement Date: 2013-05-06", y="% Return")

BHAR20130506 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/BHAR20130506.csv", header = TRUE, stringsAsFactors = FALSE)
BHAR20130506$DATE <- as.Date(as.character(BHAR20130506$DATE), format="%Y%m%d")
#BHAR20130506$DATE <- as.character(BHAR20130506$DATE)
BHAR20130506$BHAR <- (exp(cumsum(log(1+BHAR20130506$ARET)))) - (exp(cumsum(log(1+BHAR20130506$mret))))

BHAR20130506G <- ggplot(BHAR20130506) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=BHAR), group=1, color = 'red') +
  labs(title = "ANC: 36-Month Buy-and-Hold Abnormal Returns (BHARs)", subtitle="Market return (Black), NuVasive return, (Blue), NuVasive 36-Mon BHARs (Red)",
       x ="Acquisition Announcement Date: 2013-05-06", y="% Return")

Sixth <- grid.arrange(ACAR20130506G, BHAR20130506G, ncol=1)
```

**Ellipse Technologies Acquisition**

```{r, echo = FALSE, warnings = FALSE}
library(knitr)
library(ggplot2)
library(gridExtra)

ACAR20160105 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/ACAR20160105.csv", header = TRUE, stringsAsFactors = FALSE)
ACAR20160105$DATE <- as.Date(as.character(ACAR20160105$DATE), format="%Y%m%d")
ACAR20160105$DATE <- as.character(ACAR20160105$DATE)

ACAR20160105G <- ggplot(ACAR20160105) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=AbnRet), group=1, color = 'red') +
  labs(title = "Ellipse Technologies: 4-Day Cumulative Abnormal Returns (CARs)", subtitle="Market return (Black), NuVasive return Blue), NuVasive 5-Day CARs (Red)",
       x ="Acquisition Announcement Date: 2016-01-05", y="% Return")

BHAR20160105 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/BHAR20160105.csv", header = TRUE, stringsAsFactors = FALSE)
BHAR20160105$DATE <- as.Date(as.character(BHAR20160105$DATE), format="%Y%m%d")
#BHAR20160105$DATE <- as.character(BHAR20160105$DATE)
BHAR20160105$BHAR <- (exp(cumsum(log(1+BHAR20160105$ARET)))) - (exp(cumsum(log(1+BHAR20160105$mret))))

BHAR20160105G <- ggplot(BHAR20160105) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=BHAR), group=1, color = 'red') +
  labs(title = "Ellipse Technologies: 12-Month Buy-and-Hold Abnormal Returns (BHARs)", subtitle="Market return (Black), NuVasive return, (Blue), NuVasive 36-Mon BHARs (Red)",
       x ="Acquisition Announcement Date: 2016-01-05", y="% Return")

Seventh <- grid.arrange(ACAR20160105G, BHAR20160105G, ncol=1)
```

**Biotronic NeuroNetwork Acquisition**

```{r, echo = FALSE, warnings = FALSE}
library(knitr)
library(ggplot2)
library(gridExtra)

ACAR20160606 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/ACAR20160606.csv", header = TRUE, stringsAsFactors = FALSE)
ACAR20160606$DATE <- as.Date(as.character(ACAR20160606$DATE), format="%Y%m%d")
ACAR20160606$DATE <- as.character(ACAR20160606$DATE)

Eighth <- ggplot(ACAR20160606) +
  geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
  geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
  geom_line(aes(x=DATE, y=AbnRet), group=1, color = 'red') +
  labs(title = "Biotronic NeuroNetwork: 4-Day Cumulative Abnormal Returns (CARs)", subtitle="Market return (Black), NuVasive return Blue), NuVasive 5-Day CARs (Red)",
       x ="Acquisition Announcement Date: 2016-06-06", y="% Return")

# BHAR20160606 <- read.csv("/Users/gregoryrobinson/Desktop/ Greg Mac/Jobs/NuVasive/GraphData/BHAR20160606.csv", header = TRUE, stringsAsFactors = FALSE)
# BHAR20160606$DATE <- as.Date(as.character(BHAR20160606$DATE), format="%Y%m%d")
# #BHAR20160606$DATE <- as.character(BHAR20160606$DATE)
# BHAR20160606$BHAR <- (exp(cumsum(log(1+BHAR20160606$ARET)))) - (exp(cumsum(log(1+BHAR20160606$mret))))
# 
# BHAR20160606G <- ggplot(BHAR20160606) +
#   geom_line(aes(x=DATE, y=mret), group=1, color = 'black') +
#   geom_line(aes(x=DATE, y=ARET), group=1, color = 'blue') +
#   geom_line(aes(x=DATE, y=BHAR), group=1, color = 'red') +
#   labs(title = "Biotronic NeuroNetwork: 12-Month Buy-and-Hold Abnormal Returns (BHARs)", subtitle="Market return (Black), NuVasive return, (Blue), NuVasive 36-Mon BHARs (Red)",
#        x ="Acquisition Announcement Date: 2016-06-06", y="% Return")

print(Eighth)
```

Note: insufficient data for BHAR analysis

####DATA

Mergers and Acquisitions from Thomson Reuters SDC Platinum Database

* U.S. target and acquiring firms only
* Date announced and date effective between May 13, 2004 and December 31, 2016
* Deal types:
* Disclosed Value Mergers & Acquisitions
* Undisclosed Value Mergers & Acquisitions
* Tender Offers
* Exchange Offers
* Recapitalizations
* For ‘All Firms,’ ‘All SIC code 3841,’ and ‘Primary SIC code 3841 only,’ restrict to target firms with valid PERMNO 

Daily stock file, daily stock index (equal and value-weighted market portfolio returns including distributions), monthly stock file, and monthly stock index (equal and value-weighted market portfolio returns including distributions) from the Center for Research in Security Prices (CRSP) accessed via the Wharton Research Data Services (WRDS)

* Require at least 150 return observations in estimation period [-242, 2] data 
* Fama-French 4-Factor asset pricing model
* Monthly risk free rate
* Monthly portfolio returns


#### PROGRAMS USED

* SAS (remote WRDS access, create dataset)
* Stata (analyze dataset, run regressions)
* R (visualizations using ggplot2)
* Excel (dataset export)

#### REFERENCES

* SDC Platinum. (2017) SDC Platinum. Thomson Reuters. [Online]. Available at: Subscription Service (Accessed: September 2017).  

* CRSP/Compustat Merged. (2017) CRSP/Compustat Merged. Center for Research in Security Prices. [Online]. Available at WRDS: http://wrds-web.wharton.upenn.edu/wrds/ (Accessed: September 2017)

* French, Ken.  “Current Research Returns Data.” Personal Website.  Dartmouth University, Tuck School of Business.  Web.  http://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html#Research

